---
title: "P8105_hw3_tj2383"
author: "Tanvi Jain"
date: "10/4/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

##Problem 1

In this code chunk, I did some data cleaning:
-formatted the data to use appropriate variable names

-focused on the “Overall Health” topic

-included only responses from “Excellent” to “Poor”

-organized responses as a factor taking levels ordered from “Excellent” to “Poor”
```{r}
library(p8105.datasets)
data("brfss_smart2010")

brfss_data = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  filter(response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor") %>% 
  mutate(response = factor(response, levels = ordered("Excellent", "Very good", "Good", "Fair", "Poor")))
```

In this code chunk I answer questions about the cleaned brfss_data
```{r}
brfss_data %>% 
  select(year, locationabbr, locationdesc) %>% 
  filter(year == "2002") %>% 
  distinct(locationdesc, locationabbr) %>% 
  group_by(locationabbr) %>% 
  count() %>% 
  filter(n == "7")
```

In 2002, the states observed with 7 locations are CT, FL, and NC.


